Bee brain size is positively associated with urban lifestyles but not with positive population trends

### 2022-10-19 Manuscript/Summary/Summary.Rmd
edits

## Land use classification

**1) Europe**

Link with land European use classes

https://land.copernicus.eu/user-corner/technical-library/corine-land-cover-nomenclature-guidelines/html/index-clc-512.html

@@

Then, we calculate preferences and model the data.

```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("Europe")

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe")

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe")

library(patchwork)
p1 / p2 / p3 & plot_layout(guides = "collect")

Now we make more strict land use classifications

@@

And we model the data with this categories:

```{r Plotting Europe subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}

long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_subset.csv")

@@

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("Europe")

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe")

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe")

library(patchwork)
p1 + p2 + p3 + plot_layout(ncol = 1) + plot_layout(guides = "collect")

@@

**2) USA**

@@

United States land use classification with total number of occurrences within cover type in our dataset

@@

Then, we calculate preferences and model the data.

```{r Plotting USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA")

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA")

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA")

library(patchwork)
p1 / p2 / p3 & plot_layout(guides = "collect")

Note: The models have been run with non-informative priors from the brms function but some work on priors can be done... I have explored a bit and the results were quite similar overall but reviewers seem to don't like this approach as one of the goals of Bayesian is to use the prior knowledge of our variables to improve our predictions.

Now we make more strict categories of what is Natural, Urban and Agricultural. See below:

@@

And we model the data with this categories:

```{r Plotting USA subset, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width=6}

long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_subset.csv")

@@

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA")

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA")

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA")

library(patchwork)
p1 + p2 + p3 + plot_layout(ncol = 1) + plot_layout(guides = "collect")

@@

**2) Finally, we consider Europe and USA together**

```{r Plotting Europe USA, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, out.width="70%"}

data <- read_csv("Data/Europe_USA/data_preference_brain_weight_europe_usa.csv")

model1 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa.csv")

model2 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa.csv")

ggplot(model1, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe USA occurrences")

ggplot(model2, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe USA occurrences")

@@

>>>>

# Europe (4 levels)

@@

# Europe (3 levels)

@@

\clearpage

```{r Plotting Europe, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}

library(tidyverse)

model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("Europe 4 levels")+
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe 4 levels")+
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe 4 levels")  +
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("Europe 4 levels")  +
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

#Read data with land use filtered
long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_subset.csv")

@@

model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe_subset.csv")

pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("Europe 3 levels")+
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))+
    theme(legend.position = "none")

pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("Europe 3 levels")+
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))+
    theme(legend.position = "none")

pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("Europe 3 levels") +
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
    theme(legend.position = "none")

pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("Europe 3 levels") +
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
    theme(legend.position = "none")

library(patchwork)
p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")

@@

# Europe

Qualitative analysis on the statistical evidences for preference and avoidance by habitat

```{r Plotting Europe qualitative analysis, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="80%"}

library(tidyverse)

long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_qualitative.csv")

model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe_qualitative.csv")

model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe_qualitative.csv")

model3 = read_csv("Data/Europe_data/model_output_preference_it_europe_qualitative.csv")

model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe_qualitative.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA-Europe 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA-Europe 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA-Europe 4 levels")  + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("USA-Europe 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

#Load land use filtered data and models
long_data <- read_csv("Data/Europe_data/data_preference_residuals_europe_qualitative_subset.csv")

model1 = read_csv("Data/Europe_data/model_output_preference_residuals_europe_qualitative_subset.csv")

model2 = read_csv("Data/Europe_data/model_output_preference_brain_weight_europe_qualitative_subset.csv")

model3 = read_csv("Data/Europe_data/model_output_preference_it_europe_qualitative_subset.csv")

model4 = read_csv("Data/Europe_data/model_output_preference_brain_it_europe_qualitative_subset.csv")

pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA-Europe 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA-Europe 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA-Europe 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight /` IT") +
ggtitle("USA-Europe 3 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

library(patchwork)
p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")

# USA (4 levels)

@@

@@

# USA (3 levels)

@@

# USA (Gbif and MA thesis data)

```{r Plotting USA gbif and MA thesis data, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}

library(tidyverse)

model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA 4 levels")  + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

#Load land use filtered data and models
long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_subset.csv")

@@

model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_subset.csv")

pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight /` IT") +
ggtitle("USA 3 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

library(patchwork)
p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")

@@

# USA (Gbif and MA thesis data)

Qualitative analysis on the statistical evidences for preference and avoidance by habitat

```{r Plotting USA qualitative, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="80%"}

library(tidyverse)

long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_qualitative.csv")

model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_qualitative.csv")

model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_qualitative.csv")

model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_qualitative.csv")

model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_qualitative.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA 4 levels")  + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

#Load land use filtered data and models
long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_qualitative_subset.csv")

model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_qualitative_subset.csv")

model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_qualitative_subset.csv")

model3 = read_csv( "Data/Usa_data/model_output_preference_it_usa_qualitative_subset.csv")

model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_qualitative_subset.csv")

pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight /` IT") +
ggtitle("USA levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

library(patchwork)
p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")

@@

# USA (just Gbif without MA thesis data)

```{r Plotting USA gbif, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}

library(tidyverse)

long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_gbif.csv")

model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_gbif.csv")

model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_gbif.csv")

model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_gbif.csv")

model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_gbif.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA 4 levels")  + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight / IT") +
ggtitle("USA 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

#Load land use filtered data and models
long_data <- read_csv("Data/Usa_data/data_preference_residuals_usa_subset_gbif.csv")

model1 = read_csv("Data/Usa_data/model_output_preference_residuals_usa_subset_gbif.csv")

model2 = read_csv("Data/Usa_data/model_output_preference_brain_weight_usa_subset_gbif.csv")

model3 = read_csv("Data/Usa_data/model_output_preference_it_usa_subset_gbif.csv")

model4 = read_csv("Data/Usa_data/model_output_preference_brain_it_usa_subset_gbif.csv")

pp1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA 3 levels") + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA 3 levels")  + scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

pp4 = ggplot(model4, aes(x = brain_it, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = brain_it, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight /` IT") +
ggtitle("USA 3 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4")) +
theme(legend.position = "none")

library(patchwork)
p1 + pp1 + p2 + pp2 + p3 + pp3 + p4 + pp4 + plot_layout(ncol = 2) + plot_layout(guides = "collect")

# USA and Europe together

```{r Plotting USA Europe together, eval = TRUE, fig.align = 'center', echo = FALSE, warning = FALSE, message = FALSE, fig.height=10, fig.width=7, fig.pos="center", out.width="90%"}

library(tidyverse)

long_data <- read_csv("Data/Europe_USA/data_preference_residuals_europe_usa.csv")

model1 = read_csv("Data/Europe_USA/model_output_preference_residuals_europe_usa.csv")

model2 = read_csv("Data/Europe_USA/model_output_preference_brain_weight_europe_usa.csv")

model3 = read_csv("Data/Europe_USA/model_output_preference_it_europe_usa.csv")

model4 = read_csv("Data/Europe_USA/model_output_preference_brain_it_europe_usa.csv")

p1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Residuals") + 
ggtitle("USA-Europe 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Brain weight") +
ggtitle("USA-Europe 4 levels") + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))

p3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21) +
geom_line(aes(color=Habitat)) +
theme_bw() +
ylab("Habitat preference") +
xlab("Intertegular distance") +
ggtitle("USA-Europe 4 levels")  + 
scale_color_manual(values=c('Agricultural'='brown', 'Natural'='forestgreen', 'Urban'='purple4', "Seminatural"="cyan4"))
